---
title: "More Bayes"
subtitle: "And multilevel binomial logistic regression"
author: "Daniel Anderson "
date: "Week 8"
output:
  xaringan::moon_reader:
    css: ["default", "new.css"]
    lib_dir: libs
    nature:
      navigation:
        scroll: false
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
    includes:
      in_header: "load-feather.html"
---

```{r include = FALSE}
source(here::here("static", "slides", "slide-setup.R"))
xaringanExtra::use_clipboard()
library(tidyverse)
theme_set(theme_minimal(25))
knitr::opts_chunk$set(fig.width = 13, message = FALSE)
update_geom_defaults('path', list(size = 2))
update_geom_defaults('smooth', list(size = 2))
update_geom_defaults('point', list(size = 5))
update_geom_defaults('text', list(size = 9))
```

`r setup("w9p1")`

---
# Agenda
* Missing values can be treated as unknown variables (parameters) in the model, and imputed from the posterior

* Fitting multilevel logistic regression models with **brms**

* Plotting **brms** fits


---
# Popular data

```{r }
popular <- read_csv(here::here("data", "popularity.csv"))
popular
```

---
```{r include = FALSE}
library(brms)
m_brms <- brm(popular ~ extrav + (extrav|class), popular)
m_brms
```

```{r eval = FALSE}
m_brms <- brm(popular ~ extrav + (extrav|class), popular)
```

```{r echo = FALSE}
m_brms
```


---
# Plotting the Bayes fit

```{r fig.height = 5}
tibble(extrav = 1:10, class = 0) %>% 
  tidybayes::add_fitted_draws(m_brms, 
                              allow_new_levels = TRUE, 
                              n = 100) %>% 
  ggplot(aes(extrav, .value)) +
  geom_line(aes(group = .draw), size = 0.1)
```

---
# Add in raw data
```{r fig.height = 5}
library(tidybayes)
tibble(extrav = 1:10, class = 0) %>% 
  add_fitted_draws(m_brms, allow_new_levels = TRUE, n = 100) %>% 
  ggplot(aes(extrav, .value)) +
  geom_point(aes(extrav, popular), data = popular,
             color = "gray70") + #<<
  geom_line(aes(group = .draw), size = 0.1)
```


---
# New data
Lung cancer data: Patients nested in doctors

```{r }
hdp <- read_csv("https://stats.idre.ucla.edu/stat/data/hdp.csv") %>% 
  janitor::clean_names() %>% 
  select(did, tumorsize, pain, lungcapacity, age, remission)
hdp
```

---
# Predict remission
Build a model where age, lung capacity, and tumor size predict whether or not the patient was in remission. Allow the intercept to vary by the doctor ID. Fit the model using **brms**

`r countdown::countdown(3)`

--
```{r }
lc <- brm(remission ~ age*tumorsize + lungcapacity + (1|did),
          data = hdp,
          family = bernoulli(link = "logit"),
          cores = 4,
          backend = "cmdstan")
```


---
# Model summary
```{r }
summary(lc)
```

---
# Posterior predictive check
```{r }
pp_check(lc, type = "bars")
```

---
# Chains
```{r fig.height = 9}
plot(lc)
```


---
# Marginal predictions: Age
```{r }
conditional_effects(lc, "age")
```

---
# Marginal predictions: tumor size
```{r }
conditional_effects(lc, "tumorsize")
```

---
# Marginal predictions: lung capacity
```{r }
conditional_effects(lc, "lungcapacity")
```

---
# Interaction

```{r }
conditional_effects(lc, "age:tumorsize")
```

---
# Make predictions

Check the relation for tumor size

```{r }
library(tidybayes)
pred_tumor <- expand.grid(
    age = 20:80,
    lungcapacity = mean(hdp$lungcapacity),
    tumorsize = 30:120,
    did = -999
  ) %>% 
  add_fitted_draws(model = lc, n = 100,
                   allow_new_levels = TRUE)
pred_tumor
```

---
# Plot
```{r }
ggplot(pred_tumor, aes(age, .value)) +
  stat_lineribbon()
```

---
# Different plot

```{r fig.height = 5}
pred_tumor %>% 
  filter(tumorsize %in% c(30, 60, 90, 120)) %>% 
ggplot(aes(age, .value)) +
  geom_line(aes(group = .draw), alpha = 0.2) +
  facet_wrap(~tumorsize) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

---
# Variance by Doctor
Let's look at the relation between age and proability of remission for each of the first nine doctors.

--

```{r}
pred_age_doctor <- expand.grid(
    did = unique(hdp$did)[1:9],
    age = 20:80,
    tumorsize = mean(hdp$tumorsize),
    lungcapacity = mean(hdp$lungcapacity)
  ) %>% 
  add_fitted_draws(model = lc, n = 100)
```

---
```{r }
pred_age_doctor
```

---
```{r fig.height = 9}
ggplot(pred_age_doctor, aes(age, .value)) +
  geom_line(aes(group = .draw), alpha = 0.2) +
  facet_wrap(~did)
```

---
# Going further
* We can pull lots of different things from our model

--
* Let's start by looking at what's actually in the model


--
In this case `r_*` implies "random". These are the deviations from the average.

```{r }
get_variables(lc)
```

---
# Get all draws
Let's look at the intercept

```{r }
int <- lc %>% 
  spread_draws(b_Intercept)
int
```

---
# Plot the distribution

```{r }
ggplot(int, aes(b_Intercept)) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  geom_vline(xintercept = median(int$b_Intercept),
             color = "magenta",
             size = 2)
```

---
# Grab random effects

* The random effect name is `r_did`

* We use brackets to assign new names

```{r }
spread_draws(lc, r_did[did, term])
```

---
# Look at did distributions
First 80 doctors
```{r eval = FALSE}
dids <- spread_draws(lc, r_did[did, ]) # all terms, which is just one
dids %>% 
  filter(did %in% 1:80) %>% 
  ggplot(aes(x = r_did, y = factor(did))) +
  ggridges::geom_density_ridges(color = NA, fill = "#61adff") +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_blank())
```


---
```{r echo = FALSE, fig.height=11}
dids <- spread_draws(lc, r_did[did, ]) # all terms, which is just one
dids %>% 
  filter(did %in% 1:80) %>% 
  ggplot(aes(x = r_did, y = factor(did))) +
  ggridges::geom_density_ridges(color = NA, fill = "#61adff") +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_blank())
```

---
# Long format

Use `gather_draws()` to return a long format, suitable for plotting (and many other things)

```{r }
fixed_l <- lc %>% 
  gather_draws(b_Intercept, b_age, b_tumorsize, b_lungcapacity, 
               `b_age:tumorsize`)
fixed_l
```

---
# Plot the densities

```{r }
ggplot(fixed_l, aes(.value)) +
  geom_density(fill = "#61adff", alpha = 0.7, color = NA) + 
  facet_wrap(~.variable, scales = "free")
```



---
# Multiple comparisons
One of the nicest things about Bayes is that any comparison you want to make can be made without jumping through a lot of additional hoops (e.g., adjusting $\alpha$).


--
### Scenario
Imagine a **35** year old has a tumor measuring **58 millimeters** and a lung capacity rating of **0.81**.


--
What would we estimate as the odds of remission if this patient had `did == 1` versus `did == 2`?


---
# Fixed effects

Not really "fixed", but rather just average relation

```{r }
fe <- lc %>% 
  spread_draws(b_Intercept, b_age, b_tumorsize, b_lungcapacity, 
               `b_age:tumorsize`)
fe
```

---
# Data
```{r }
age <- 35
tumor_size <- 58
lung_cap <- 0.81
```

--
population-level predictions

```{r }
pop_level <- 
  fe$b_Intercept +
  (fe$b_age * age) +
  (fe$b_tumorsize * tumor_size) +
  (fe$b_lungcapacity * lung_cap) +
  (fe$`b_age:tumorsize` * (age * tumor_size))
pop_level
```

---
# Plot

```{r fig.height = 6}
pd <- tibble(population_level = pop_level) 

ggplot(pd, aes(population_level)) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  geom_vline(xintercept = median(pd$population_level),
             color = "magenta",
             size = 2)
```

---
# Add in did estimates

```{r }
did1 <- filter(dids, did == 1)
did2 <- filter(dids, did == 2)

pred_did1 <- pop_level + did1$r_did
pred_did2 <- pop_level + did2$r_did
```

---
# Distributions

```{r }
did12 <- tibble(did = rep(1:2, each = length(pred_did1)),
                pred = c(pred_did1, pred_did2))

did12_medians <- did12 %>% 
  group_by(did) %>% 
  summarize(did_median = median(pred))

did12_medians
```

---
# Plot
```{r fig.height = 6}
ggplot(did12, aes(pred)) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  geom_vline(aes(xintercept = did_median), data = did12_medians,
             color = "magenta",
             size = 2) +
  facet_wrap(~did, ncol = 1)
```

---
# Transform
Let's look at this again on the probability scale. Note I'm using `brms::inv_logit_scaled()` to make the transformation.

--
```{r eval = FALSE}
ggplot(did12, aes(inv_logit_scaled(pred))) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  geom_vline(aes(xintercept = inv_logit_scaled(did_median)), 
             data = did12_medians,
             color = "magenta",
             size = 2) +
  facet_wrap(~did, ncol = 1)
```

---
```{r echo = FALSE}
ggplot(did12, aes(inv_logit_scaled(pred))) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  geom_vline(aes(xintercept = inv_logit_scaled(did_median)), 
             data = did12_medians,
             color = "magenta",
             size = 2) +
  facet_wrap(~did, ncol = 1)
```

---
# Difference
* The difference in the probability of remission for our theoretical patient is large between the two doctors.

* The median difference in log-odds is

```{r }
diff(did12_medians$did_median)
```

so the patient is about 3.5 times **more likely** to have their cancer go into remission if they had did 2, instead of 1.

--
## How confident are we in this difference?

---
# Everything is a distribution
Just compute the difference in these distributions, and we get a new distribution, which we can use to summarize our uncertainty

--
```{r }
did12_wider <- did12 %>% 
  mutate(.chain = rep(rep(1:4, each = 1000), 2),
         .draw = rep(1:1000, 8)) %>% 
  pivot_wider(names_from = "did", values_from = "pred")

did12_wider
```

---
# Compute difference
```{r }
did12_wider <- did12_wider %>% 
  mutate(diff = `2` - `1`)

did12_wider
```

---
# Summarize

```{r }
quantile(did12_wider$diff, 
         probs = c(0.025, 0.5, 0.975))
```


---
# Plot distribution

```{r }
ggplot(did12_wider, aes(diff)) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  geom_vline(aes(xintercept = median(diff)), 
             color = "magenta",
             size = 2)
```

---
# Directionality
Let's say we want to simplify the question to directionality.


--
Is there a greater chance of remission for `did` 2 than 1?

--
```{r }
table(did12_wider$diff > 0) / 4000
```


--
The distributions are not overlapping at all - therefore, we are as certain as we can be that the odds of remission are higher with `did` 2 than 1.


---
# One more quick example
Let's do the same thing, but comparing `did` 2 and 3.

```{r }
did3 <- filter(dids, did == 3)
pred_did3 <- pop_level + did3$r_did

did23 <- did12_wider %>% 
  select(-`1`, -diff) %>% 
  mutate(`3` = pred_did3,
         diff = `3` - `2`)
did23
```

---
# Directionality

```{r }
table(did23$diff > 0) / 4000
```

So there's roughly an 87% chance that the odds of remission are higher with with `did` 3 than 2.

---
# Plot data

```{r }
pd23 <- did23 %>% 
  pivot_longer(`2`:diff, 
               names_to = "Distribution",
               values_to = "Log-Odds")
pd23
```

---
```{r }
ggplot(pd23, aes(`Log-Odds`)) +
  geom_histogram(fill = "#61adff",
                 color = "white") +
  facet_wrap(~Distribution, ncol = 1)
```

---
class: inverse-green middle
# Next time
## Growth Modeling 2
We'll continue to discuss and use Bayes
